---
title: "Project 1 - knitr in data analysis"
author: "Reut Timor"
date: "October 30, 2017"
output: 
  html_document: 
    keep_md: yes
---
The data for this assignment can be downloaded from the course web site:

Dataset: [Activity monitoring data] (https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

turn warnings off 

```{r}
knitr::opts_chunk$set(warning=FALSE)
```


##1. Code for reading in the dataset and preprocessing the data


```{r}
library(ggplot2)
library(data.table)
url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,"activity.zip")
unzip("activity.zip")
data_activ <- read.csv("activity.csv")
Sys.setenv(TZ="Europe/Berlin")
data_activ$date <- as.POSIXct(data_activ$date, "%Y-%m-%d")
weekday <- weekdays(data_activ$date)
data_activ <- cbind(data_activ,weekday)

summary(data_activ)
```


##2. Code for creating a plot presenting the mean steps taken each day

```{r}
sumsteps <- with(data_activ, aggregate(steps,list(date),sum, na.rm = TRUE))
names(sumsteps) <- c("date", "steps")
hist(sumsteps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day")
stepsmean<-mean(sumsteps$steps,na.rm = TRUE)
stepsmedian<-median(sumsteps$steps)
```


3. The mean of all steps is `r stepsmean` 
The median is `r stepsmedian`


##4. Code for tracking the time series plot of the average number of steps taken


```{r}
average_data_activ <- with(data_activ,aggregate(steps, list(interval), mean, na.rm=TRUE))
names(average_data_activ) <- c("interval", "mean")
plot(average_data_activ$interval, average_data_activ$mean, type = "l", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```


##5. The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps is:


```{r}
average_data_activ[which.max(average_data_activ$mean), ]$interval
```


##6. Code to describe and show a strategy for imputing missing data


Since there are missing values in the data, filling those gaps should ease calculations on the data. 


```{r}
Na_data<-is.na(data_activ$steps)
```
## The number of missing values is `r length(Na_data)`


One approch to fill missing values instead of ignoring them is to use the mean value of the 5 min interval


```{r}
imputed_steps <- average_data_activ$mean[match(data_activ$interval, average_data_activ$interval)]
```


Now, lets create a new dataset with imputed values instead of the Na's

```{r}
data_activ_imputed <- transform(data_activ, steps = ifelse(is.na(data_activ$steps), yes = imputed_steps, no = data_activ$steps))
sumsteps_imputed <- aggregate(steps ~ date, data_activ_imputed, sum)
names(sumsteps_imputed) <- c("date", "steps")
means<-mean(sumsteps_imputed$steps)
medians<-median(sumsteps_imputed$steps)

```


##7. Histogram of the total number of steps taken each day after missing values are imputed


```{r}
hist(sumsteps_imputed$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day")
```

The mean of all steps after imputed values is `r means` 
The median of all steps after imputed values is `r medians`

##8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


First, a new factor variable in the dataset with two levels - "weekday" and "weekend"  will indicate if a given date is a weekday or weekend day.


```{r}
data_activ$date <- as.Date(strptime(data_activ$date, format="%Y-%m-%d"))
data_activ$datetype <- sapply(data_activ$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```


A panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 


```{r}
data_activ_date <- with(data_activ,aggregate(steps~interval + datetype, data_activ, mean, na.rm = TRUE))
plot<- ggplot(data_activ_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```